Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-04-20 CONCUR2018/paper.tex
revised experimental section

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{p} & \multicolumn{1}{|c|}{No.} & \multicolumn{4}{|c|}{CEGIS-AR} & \multicolumn{4}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found}\\\hline
1  & DC Motor          & 2 & 100 & 1.3\,s & 26.9\,s& 22& 20& 1.1\,s &   3.8\,s & 2 & 58\\
2  & Helicopter        & 3  & 100 & 57.2\,s & t.o. & 54 & 15 & 7.1\,s  &   18.0\,s & 9 & 90\\
3  & Inverted Pendulum & 4 & 100 &  277\,s& t.o. & 82& 0 & 20.0\,s  &  t.o. & 55 & 20\\
4  & Magnetic Pointer  & 2  & 100 & 5.7\,s & 229.9\,s & 38 & 42 & 3.0\,s  &  9.5\,s & 5 & 95 \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 0.8\,s & 53.6\,s & 27 & 71 &0.4\,s  &   4.9\,s & 6 & 92\\
6  & Suspension        & 4 & 100 &15.0\,s & 177.2\,s & 30 & 70 & 14.8\,s&166.9\,s  & 26 & 73\\
7  & Tape Driver       & 3 & 100 & 4.0\,s& 17.7\,s & 11 & 80 &3.1\,s  &   16.8\,s & 8 & 82\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results.  Minimum, and Median times for standard CEGIS-AR (with
abstraction refinement~\cite{DBLP:conf/cav/AbateBCCDKKP17}) vs CEGIS-OR on
multiple benchmarks.  No.  indicates Number of benchmarks, t.o.~indicates
timeouts (in the median case, it indicates more than half of the benchmarks
timed out).  p denotes the number of variables of a model.}
\end{table}

Using the described technique we managed to synthesize controllers for $509$
benchmarks (compared to $290$ found by our baseline), with a median time of
$16.8$ seconds.
%
%and the maximum observed synthesis time was under 10 minutes for a $4^{th}$ order system ($8^{th}$ order considering the observer).
We consider these times short enough to be of practical use to control
engineers.  Regarding the cases for which we fail to find a controller, we
note that we use a relatively short timeout (5 minutes), so it is possible
that longer timeouts would instead lead to a valid solution.  However, there
are cases where a solution may not be found due the approximations in the
synthesizer that would prevent a true candidate from being found or
over-approximations in the verifier that may reject a valid candidate.  It
is impossible for us to differentiate these from case where a controller
does not exist.

We found that on average, observer synthesis is over an order of magnitude slower
than full state controller synthesis for the given benchmarks (SAT has
a worst case exponential time complexity, and typical polynomial time, so higher
order systems will show larger differences).
In order to show the improvements due to our optimization-enhanced CEGIS, we
compared our approach against a standard CEGIS with abstraction refinement
procedure~\cite{DBLP:conf/cav/AbateBCCDKKP17} (We note that the naive approach
in~\cite{DBLP:conf/cav/AbateBCCDKKP17} cannot handle the continuous time case.  
Furthermore, since it is in general slower than the abstraction refinement
version, the latter is a more representative benchmark).
For the majority of our benchmarks, we observe that our approach is
significantly faster than~\cite{DBLP:conf/cav/AbateBCCDKKP17}. 
We further note that the number of times CEGIS times out is much larger than
CEGIS-OR which results in more found solutions.

@@

Overall, we observed more improvements with CEGIS-OR for
systems where the safety specification is tight, 
meaning that the standard CEGIS loop would require many iterations.
%  We obtain the most improvements for the models with large orders?}
%More interesting cases appear when we increase the order, showing
%% For instance, benchmark X, which timed-out when using standard CEGIS,
%% is running under 2 minutes using CEGIS-OR. This is the result of
%% removing several multiplications which are known to be difficult for
%% SAT solvers \note{I'm not sure readers will relate to where this is happening in the paper. is it the fact that you use that canonical form? then say that. however, i would have thought you do that in
%% both approaches..}.
These improvements are due to the Abstraction Refinement and Optimization in CEGIS-OR.
While abstraction refinement enables us to start with coarse abstractions
that are only refined on demand, 
the optimization
allows to locally improve candidate solutions and avoid entire iterations of the CEGIS loop
(which would have also required more refined abstractions).
%% elimination of counterexamples that
%% would have been created without the refinement. Since further
%% synthesis require increasingly complex models which are linear in the
%% number of multiplications, eliminating these results in large
%% improvements.
While some systems might still require a very precise
abstraction resulting in many refinements
(this is generally the cause of time-outs in CEGIS-OR),
others allow us to find solutions with rather coarse
abstractions (this is reflected in the smaller minimum
times obtained with CEGIS-OR). 
%Since further synthesis require increasingly
%complex models which are linear in the number of multiplications,
%eliminating  these results in large improvements.  In cases when the initial candidate is
%suitable, we gain nothing with the refinement and therefore find similar times.

>>>>

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{p} & \multicolumn{1}{|c|}{No.} & \multicolumn{4}{|c|}{CEGIS\cite{DBLP:conf/cav/AbateBCCDKKP17}} & \multicolumn{4}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found}\\\hline
1  & DC Motor          & 2 & 100 & 1.3\,s & 26.9\,s& 22& 20& 1.1\,s &   3.8\,s & 2 & 58\\
2  & Helicopter        & 3  & 100 & 57.2\,s & t.o. & 54 & 15 & 7.1\,s  &   18.0\,s & 9 & 90\\
3  & Inverted Pendulum & 4 & 100 &  277\,s& t.o. & 82& 0 & 20.0\,s  &  t.o. & 55 & 20\\
4  & Magnetic Pointer  & 2  & 100 & 5.7\,s & 229.9\,s & 38 & 42 & 3.0\,s  &  9.5\,s & 5 & 95 \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 0.8\,s & 53.6\,s & 27 & 71 &0.4\,s  &   4.9\,s & 6 & 92\\
6  & Suspension        & 4 & 100 &15.0\,s & 177.2\,s & 30 & 70 & 14.8\,s&166.9\,s  & 26 & 73\\
7  & Tape Driver       & 3 & 100 & 4.0\,s& 17.7\,s & 11 & 80 &3.1\,s  &   16.8\,s & 8 & 82\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
%\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results.  Minimum, and Median times for standard CEGIS-AR (with
abstraction refinement~\cite{DBLP:conf/cav/AbateBCCDKKP17}) vs CEGIS-OR on
multiple benchmarks.  No.  indicates Number of benchmarks, t.o.~indicates
timeouts (in the median case, it indicates more than half of the benchmarks
timed out).  p denotes the number of variables of a model.}
\end{table}

Using the described technique we managed to synthesize controllers for $509$
benchmarks (compared to $290$ found by our baseline), with a median time of
$16.8$ seconds.
%
%and the maximum observed synthesis time was under 10 minutes for a $4^{th}$ order system ($8^{th}$ order considering the observer).
We consider these times short enough to be of practical use to control
engineers.  Regarding the cases for which we fail to find a controller, we
note that we use a relatively short timeout (5 minutes), so it is possible
that longer timeouts would instead lead to a valid solution.  However, there
are cases where a solution may not be found due the approximations in the
synthesizer that would prevent a true candidate from being found or
over-approximations in the verifier that may reject a valid candidate.  It
is impossible for us to differentiate these from the case where a controller
does not exist.

In order to show the improvements due to our optimization-enhanced
CEGIS, we compared our approach against the CEGIS approach with
abstraction refinement described
in~\cite{DBLP:conf/cav/AbateBCCDKKP17}.  We note that,
while~\cite{DBLP:conf/cav/AbateBCCDKKP17} presents two approaches, one
of them (referred to as the naive approach
in~\cite{DBLP:conf/cav/AbateBCCDKKP17}) cannot handle the continuous
time case and thus we can only compare against the abstraction-based
one (referred to as CEGIS~\cite{DBLP:conf/cav/AbateBCCDKKP17} in Table~\ref{tab:cegis_results}).  %% Furthermore, since it is in general slower than the abstraction
%% refinement version, the latter is a more representative benchmark).
For the majority of our benchmarks, we observe that our approach is
significantly faster than CEGIS~\cite{DBLP:conf/cav/AbateBCCDKKP17}.  %% We
%% further note that the number of times CEGIS times out is much larger
%% than CEGIS-OR which results in more found solutions.

@@

Overall, we observed more improvements with CEGIS-OR for
systems where the safety specification is tight, 
meaning that the CEGIS~\cite{DBLP:conf/cav/AbateBCCDKKP17}
loop would require many iterations.
%  We obtain the most improvements for the models with large orders?}
%More interesting cases appear when we increase the order, showing
%% For instance, benchmark X, which timed-out when using standard CEGIS,
%% is running under 2 minutes using CEGIS-OR. This is the result of
%% removing several multiplications which are known to be difficult for
%% SAT solvers \note{I'm not sure readers will relate to where this is happening in the paper. is it the fact that you use that canonical form? then say that. however, i would have thought you do that in
%% both approaches..}.
These improvements are due to the interplay between Abstraction
Refinement and Optimization in CEGIS-OR.  While abstraction refinement
enables us to start with coarse abstractions that are only refined on
demand, the optimization allows to locally improve candidate solutions
and avoid entire iterations of the
CEGIS~\cite{DBLP:conf/cav/AbateBCCDKKP17} loop (which would have also
required more refined abstractions).
%% elimination of counterexamples that
%% would have been created without the refinement. Since further
%% synthesis require increasingly complex models which are linear in the
%% number of multiplications, eliminating these results in large
%% improvements.
While some systems might still require a very precise abstraction
resulting in many refinements (this is generally the cause of
time-outs in CEGIS-OR), others allow us to find solutions with rather
coarse abstractions (this is reflected in the smaller minimum times
obtained with CEGIS-OR).

With respect to the additional complexity introduced by the synthesis
of observer-based controllers compared to fully observed systems, we
found that on average, observer synthesis is over an order of
magnitude slower than full state controller synthesis for the given
benchmarks. This is due to the fact that the introduction of the
observer doubles the order of the model.

%% (SAT has a worst case exponential time complexity, and
%% typical polynomial time, so higher order systems will show larger
%% differences).

%Since further synthesis require increasingly
%complex models which are linear in the number of multiplications,
%eliminating  these results in large improvements.  In cases when the initial candidate is
%suitable, we gain nothing with the refinement and therefore find similar times.
